Newly Left-Behind Children's Educational Performance Drops Slightly After Parental Migration

### 2023-11-09 MScThesisText.Rmd
Add latest literature

author: 
  - name          : "Jiancheng Gu"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Faculty of Social Sciences, Vrije Universiteit Amsterdam"

note: |
  May 15, 2023

authornote: | 
  This paper's open science repository: https://github.com/jianchenggu/MScThesis
  An earlier version of this article was submitted in partial fulfillment of the requirements for the degree -- Master of Science in Societal Resilience (Research).

@@

keywords          : "labor migration, left-behind children, educational performance"
wordcount         : "6,789"

floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no

@@

classoption       : "man"
output            : papaja::apa6_pdf
appendix: 
  - "appendix.Rmd"
  - "appendix-b.Rmd"
bibliography: references.bib
---

@@

# Introduction

@@

How are stay-behind children faring in the absence of parents? Research findings are mixed due to the differential effects of parental migration. Consider children's education. According to the review articles by @venta2022 and @vanhook2020, labor migrants send remittances back, which can fund children's schooling as well as improve their nutritional and living conditions. To what degree can the inflow of remittances compensate for the loss of parental attention? Parental migration decreases the amount and quality of parental guidance, protection, and support. It is uncertain how these two opposing effects together shape children's educational performance. This uncertainty merits empirical investigation.

A challenge arises in estimating the effect of parent's migration on children's educational achievement. Experiments can unveil causal relations. It is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, a quasi-experimental design can help with causal inference. It requires that the cause be on par with random assignment to participants, conditional on identification assumptions.

@@

## China's Context

@@

The institutional discrimination against labor migrants finds its roots in China's "multilevel citizenship" system. As @vortherms2021 explained, the constituents of citizenship -- "membership rules and rights entitlements" -- operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through *hukou*, or the household registration system:

> A political institution rooted in centuries of Chinese bureaucracy, the modern hukou is the primary identity document defining membership and entitlements for local citizenship. Accessing fundamental citizenship rights depends on your registration status ... Registration includes basic identifying information, a specific address of registration, and a type: urban, rural, or resident. [@vortherms2021]

In other words, the hukou system binds one's citizenship rights with one's place of registration. In theory, Chinese citizens can migrate freely within the country. In practice, they cannot change hukou freely. When rural residents move to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the host city [@zhang2018; @wan2022]. Migrant workers have organized petitions, strikes, and protests. These actions often fell on deaf ears and sometimes even ended with police crackdown [@chan2019; @göbel2021]. Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns. Only during the winter and summer breaks can these children reunite with parents.

How does China counter the issue of labor migration and left-behind children? According to @wang2020 and @dong2020, the government has gradually liberalized the multilevel citizenship scheme. It has introduced residence permit for internal migrants as the basis for social entitlements. First-tier cities like Beijing and Shanghai have brought forth a point-based system to grant hukou to newcomers. The system's evaluation criteria include in-demand talents and skills, educational and professional qualifications, and consecutive years of employment and tax-payment. This parallels how European states differentially incorporate foreign immigrants into national welfare regimes [@pasquali2022]. Another policy response, as stated by @roberts2018, is the centralized boarding school. The Ministry of Education promotes boarding schools in the rural county-seats. These schools serve as a cost-effective way to enroll students in surrounding villages and to look after the left-behind children. In reality, such schools "are often ill-equipped to ensure the healthy upbringing and psychological wellbeing of children" (p. 6).

Social scientists have researched extensively about the education of Chinese left-behind children, but the current literature still has three major shortcomings. The first shortcoming lies in sampling. For many studies, the geographic coverage is limited to one to a few counties; The participants are children left behind in rural areas but not those in urban areas [e.g. @hu2019; @zhou2014]. @alpermann2020 pointed that the Chinese government reinforces "an urban hierarchy" besides the rural-urban divide. In the hierarchy's bottom are small- and medium-sized cities in the inland. They also see parents migrating to metropolises and leaving children behind [@wang2023]. The National Bureau of Statistics noted that the number of urban left-behind children have been growing, and the trend is expected to continue [@nbs2017]. The second shortcoming lies in research design. Some studies rely on cross-sectional data that focus on single time points [e.g. @khalid2022]. These studies fail to capture the dynamic nature of migratory journey. Some other studies probe correlation from panel data using basic ordinary least squares regression [e.g. @chen2023a]. These studies could have employed causal inference techniques. The third shortcoming lies in the measurement of outcome. @chang2019 and @wang2019 measured students' educational performance with the standardized test score of only one academic subject, namely mathematics. To tackle these limitations, I adopted a nationally representative two-wave panel dataset. My sample includes urban left-behind children as well. I included the test results of multiple subjects to measure comprehensive academic achievement.

## Theory and Hypotheses

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital [@debrauw2019; @shen2021]. Economic capital is, in Bourdieu's words, "immediately and directly convertible into money and may be institutionalized in the form of property rights" [-@bourdieu2002, p.16]. Economic resources can help individuals acquire human capital, which is the stock of "knowledge, skills, and health that people accumulate over their lives" [@worldbank2020, p.1]. @yang2008 noted that in the Philippines, a sharp depreciation of peso increased the amount of remittances from migrants working abroad. This subsequently boosted children's school enrollment and educational spending. @vikram2021 performed propensity score matching and fixed-effect regression on a two-wave panel dataset featuring Indian children aged 8-11. She showed that fathers' migration was linked with increases in left-behind children's reading capability and educational expenditure (p. 991). In China, @gu2022a interviewed migrant workers and their family members. She found that they aspire to move up the social ladder; Such aspiration translates into "intensive and all-out investment in children's education" (p. 526). Oftentimes, labor migration presented the only way to sustain their children's education and livelihood. @peng2018 interviewed first-generation migrant mothers whose initial migration dated back to the 1990s. To them, "good mothering" primarily meant offering economic support for their children in education, which can help their children climb the ladder of success.

The other theoretical model is the family disruption model [@shen2021]. Children flourish under parental nurturing care. This term refers to the "stable environments that promote health and adequate nutrition, protect from threats ... opportunities for learning," and "responsive, emotionally supportive and developmentally enriching relationships" [@black2021, p. 1]. Parental migration decreases nurturing care, as characterized by less homework supervision and emotional support by parents. Hong talked to a group of left-behind children at a rural school [@hong2019]. These students expressed the utter sense of loneliness "with no advocate, no support and with no one looking out for them" (p. 13). Many of them felt indifferent to grades and disengaged with school. Some even longed to quit school to pursue a free, independent life (p. 14). @murphy2022 talked to left-behind adolescents at a Chinese village. Some of them "embraced the 'left-behind' label" because it captured their need for more parental care (p. 185), "especially companionship and affection in daily life" (p. 186). According to @chen2023, caregiver-child interactions are infrequent and unstable for left-behind children. She showed that the loss in such interactions "mediates almost all of the adverse effects of parental absence." Her conclusion was based on structural equation modeling using a cross-sectional survey with a sample of Chinese children aged three to fifteen.

Given the two opposing effects, researchers have obtained mixed results on how parental migration impacts children's education. @botezat2020 adopted past immigration rate as an instrumental variable. They found a positive effect of parents' migration on academic grades of adolescents aged eleven to fifteen in Romania. The team of @wang2019 collected data from junior high school students in central China. Their two-way fixed-effect model showed no significant effect of parental migration on left-behind children's achievement in mathematical test. @nguyen2016 conducted fixed-effect regression and demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight.

@@

Researchers have studied whether the impact differs between mother's migration and father's migration. @xu2019 adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had "little or no association with negative outcomes" on children's academic and cognitive performances. Only-mother absence, however, showed a "strong association with negative outcomes" (p. 1646). @chen2019 conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that maternal migration was "negatively associated with children's social competence and academic performance," while paternal migration had no direct effects on the two outcomes (pp. 860-861).

@murphy2020 noted that in some Asian and Latin American societies, the social norms designate men as breadwinners and women as caregivers. Under such norms, fathers' migration is more likely to win children's acceptance than does mothers' migration. Left-behind children may even resent their migrant mothers for transgressing on the expected care-giving roles [@murphy2020]. I propose the second hypothesis:

@@

# Methods

## Sampling

@@

## Measurement

@@

The lists of covariates are in Table \@ref(tab:tab-ctvar-dit) -- Table \@ref(tab:tab-ctvar-msr) in the Results section. At the children's level, I controlled for age, sex, self-rated health, ethnicity, hukou type, number of siblings, whether the child attends boarding school, and whether the child had skipped grade or repeated grade. Age, sex, and health are associated with educational performance [@gu2021a]. Regarding ethnicity and hukou status, Han Chinese and urban hukou holders outperform ethnic minorities and rural hukou holders at school [@hasmath2022; @xu2018; @huang2020]. I controlled for boarding at school because research results vary regarding the effect of boarding at school on academic performance [@liu2020a; @guo2021]. I also included the number of siblings as a covariate, considering the quantity-quality trade-off in multi-child family [@guo2022]. At the household level, I controlled for the levels of education of father and mother, whether the parent has a white collar job, whether the household has internet access, the perceived household financial conditions, and the number of extracurricular books at home. These variables measure family socioeconomic status, which is related to children's educational achievement [@liu2020; @young2018]. Finally, I controlled for the level of education that the child aspires to achieve and the parental expectation on the child's level of education. Children's academic achievement, their educational aspiration, and their parents' expectations are interrelated [@chen2023b; @xu2021]. All control variables were measured in the baseline survey.

## Analytical Strategy

@@

I adopted `R` (version 4.3.0) for all the analyses [@R-base]. This involved the following packages: `tidyverse` (version 2.0.0) and `haven` (version 2.5.2) for data cleaning [@R-tidyverse; @R-haven]; `fixest` (version 0.11.1) for econometric modeling [@R-fixest]; `MatchIt` (version 4.5.3) and `cobalt` (version 4.5.0) for matching [@ho2011; @R-MatchIt; @R-cobalt]; `modelsummary` (version 1.3.0), `vtable` (version 1.4.2), and `kableExtra` (version 1.3.4) for presenting the results [@R-modelsummary; @R-vtable; @R-kableExtra]; and `papaja` (version 0.1.1) for typesetting [@R-papaja].

# Results

## Descriptive Statistics

Table \@ref(tab:tab-idpvar) illustrates the independent variables, namely the migration status reported in the wave two survey. All `r printnum(nrow(d_cl))` children were living with parents in one household in baseline, but about sixteen percent saw at least one parent migrated between the two waves of survey. I call these students "newly left-behind children" and their households "new migration households." In these new migration households, only-mother migration and only-father migration each accounted for two out of five cases. The remaining one-fifth were both-parents migration.

\renewcommand{\arraystretch}{0.72}

```{r tab-idpvar}
datasummary_df(
  data = tab_idpvar,
  output = 'markdown',
  title = 'Independent variables by parental migration'
)
```

The dependent variables are summarized by the control group versus the treatment group in Table \@ref(tab:tab-dpvar) below and Figure \@ref(fig:v-plot-cog) -- Figure \@ref(fig:v-plot-ttl-1) in the appendix. Compared to non-left-behind peers, newly left-behind children scored lower in wave two's cognitive and academic tests; Their cognitive test score increased less and their academic exam score dropped more. T-tests show that for the baseline outcomes, the differences of the mean values between-group are probably due to chance. For the wave two outcomes and the between-wave changes, the differences of the mean value are statistically significant.

```{r tab-dpvar}
tab_dpvar |> 
  kbl(booktabs = T, caption = 'Dependent variables by parental migration', row.names = F) |>
  add_indent(c(2,3,5,6)) |>
  add_header_above(c(" ", "No = 3465" = 2, "Yes = 649" = 2)) |>
  add_header_above(c(" ", "Parental Migration" = 4)) |>
  footnote(general = '* p < 0.05, ** p < 0.01, *** p < 0.001')
```

Table \@ref(tab:tab-ctvar-dit) summarizes the bivariate analysis on the binary control variables. New migration households had lower internet access rate. Among the newly left-behind children, male sex, past grade retention, and boarding at school were more common compared to the non-left-behind peers. Chi-square tests show that these differences in variance are unlikely due to chance.

```{r tab-ctvar-dit}
tab_ctvar_dit |>
  kbl(booktabs = T, caption = 'Dichotomous control variables by parental migration', row.names = F) |>
  add_header_above(c(" ", "No = 3465" = 2, "Yes = 649" = 2)) |>
  add_header_above(c(" ", "Parental Migration" = 4)) |>
  footnote(general = '0 = No, 1 = Yes; * p < 0.05, ** p < 0.01, *** p < 0.001')
```

Table \@ref(tab:tab-ctvar-cnt) displays the bivariate analysis for categorical control variables, and Table \@ref(tab:tab-ctvar-msr) documents their measurement scales. While the majority of mothers and fathers had only finished high school or below, they held big expectations for children. Eight out of ten parents wished that their children complete university education. Scholars like @luo2013 and @ng2020 explained this pattern with the Chinese social norms and cultural traditions. In ancient China, the Confucian teaching emphasized the cultivation of personhood through learning, diligence, and persistence. In real life, knowledge could translate into power and privilege. Ordinary people could study for the highly competitive *keju* (civil service examination). As the world's "earliest meritocratic institution," this national exam system recruited top performers into the imperial government and the elite class [@chen2020, p. 2030]. Nowadays in China, the emperor and *keju* are long gone, but the influences are long-lasting. Chinese parents hope that their children receive the best education to secure a decent career [@luo2013; @ng2020].

```{r tab-ctvar-cnt}
tab_ctvar_cnt |>
  kbl(booktabs = T, caption = 'Categorical and continuous control variables by parental migration', row.names = F) |>
  add_header_above(c(" ", "No = 3465" = 2, "Yes = 649" = 2)) |>
  add_header_above(c(" ", "Parental Migration" = 4)) |>
  footnote(general = '* p < 0.05, ** p < 0.01, *** p < 0.001')
```

```{r tab-ctvar-msr}
datasummary_df(
  data = tab_var,
  title = 'Measurement and distribution of categorical variables',
  output = 'markdown'
)
```

## Regression Analysis

I first explain the results on children's cognitive test score. The unadjusted DID models show that newly left-behind adolescents are more likely to score lower in cognitive tests than do their non-left-behind counterparts (see Table 6 below). This lends support to Hypothesis 1a. Parental migration is linked to a `r coef(did_cog)[1] |> abs() |> round(2)` points decrease in cognitive test score. Hypothesis 2a is only partially supported. Compared with the non-left-behind peers, children with only father absent are associated with greater decrease (`r coef(did_cog_f_only)[1] |> abs() |> round(2)` points) in cognitive test score than those with only mother absent (`r coef(did_cog_m_only)[1] |> abs() |> round(2)` points). The adjusted DID models produce results similar to the unadjusted models, with one exception (see Table A1 in the appendix). That is, for the cognitive score, the both-parent-migrated coefficient's magnitude is lower than that of father-migrated.

@@

Next, I set forth the results on children's academic test grades. In line with Hypothesis 1b, the unadjusted DID models show that newly left-behind adolescents tend to perform worse in school exams than their non-left-behind peers (see Table 7 below). Parental migration is associated with a drop of `r coef(did_ttl)[1] |> abs() |> round(2)` points in academic exam score. Hypothesis 2b finds partial support. The exam grade of only-mother-migrated children sees a `r coef(did_ttl_m_only)[1] |> abs() |> round(2)` points decrease relative to non-left-behind counterparts. In the scenarios of only father migrated and both parents migrated, the results are not statistically significant. The adjusted DID models' results largely resemble those from the unadjusted models (See Table A2 in the appendix). Adding control variables measured in the baseline survey does not change the coefficients for parental migration.

@@

Some predictors are strong in both models of cognitive and academic abilities. These are children's age, preschool attendance, and educational aspiration as well as parents' expectation on children. For students, one year older in age is associated with decreases of `r coef(did_adj_cog)[3] |> abs() |> round(2)` points for cognitive test and of `r coef(did_adj_ttl)[3] |> abs() |> round(2)` points for academic exam. Preschool attendance is linked to increases of `r round(coef(did_adj_cog)[8],2)` points for cognitive test and of `r round(coef(did_adj_ttl)[8],2)` points for academic exam. Higher educational expectation from parents and greater educational aspiration from children are both linked to higher scores. The coefficients are `r round(coef(did_adj_cog)[14],2)` points and `r round(coef(did_adj_cog)[12],2)` points for cognitive test, respectively; and `r round(coef(did_adj_ttl)[14],2)` points and `r round(coef(did_adj_ttl)[12],2)` points for academic exams, respectively.

Several predictors are strong in predicting one outcome but not the other. Covariates that only relate to academic grades are children's gender and self-reported health, and the number of extracurricular books at home. Girls outperform boys by `r round(coef(did_adj_ttl)[4],2)` points in academic exams, but do not show advantage in cognitive tests. Students who consider themselves healthier are associated with higher scores in academic exams. On the contrary, those who report having more extracurricular books at home are associated with lower scores in academic exams. Covariates that only relate to cognitive scores is the number of siblings. Having an additional sibling is associated with lower cognitive score of `r coef(did_adj_cog)[7] |> abs() |> round(2)` points.

## Matching

Parental migration is not a randomly assigned treatment. Possibly, smart and ambitious students perform better at school, and high performance and ambition prompt parents to migrate to support children's schooling. Another scenario is that poverty may hinder children's educational achievement while driving parents to migrate. The endogeneity leads to selection bias. To tackle this bias, an applicable technique is matching. It aims to make two groups comparable by pairing a unit in the treatment group with one or several units in the control group based on observable covariates [@sävje2021].

A popular choice for matching, as @pan2018 introduced, is propensity score matching (PSM). For each unit in the sample, PSM will first calculate a propensity score, namely the probability of receiving treatment based on the specified covariates. This score serves as a measure of the distance between two units. When handling this distance, PSM can match the units with two approaches. One is the nearest-neighbor approach. It examines all the treated units and selects the control unit closest in distance for pairing. This approach does not try to optimize any condition; each pair of units ignores how other pairing will occur or have occurred. The second approach, the optimal-pair approach, works in a similar fashion but tries to minimize the absolute pairwise distances in total [@pan2018].

@@

@@

# Discussion and Conclusion

@@

## Limitations

@@

## Recommendations

@@

# (APPENDIX) the appendix {.unnumbered}

@@

```{r child = "appendix-b.Rmd"}
```

>>>>

author: 
  - name          : "Jiancheng Gu"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Dr.-Karl-Dorrek-Strasse 30, 3500 Krems, Austria"
    email         : "jiancheng.gu@donau-uni.ac.at"

affiliation:
  - id            : "1"
    institution   : "Department for Migration and Globalisation, Universität für Weiterbildung Krems"

note: |
  `r format(Sys.time(), '%B %d, %Y')`

authornote: | 
  This paper's open science repository: https://github.com/jchgu/MScThesis

@@

keywords          : "labor migration, left-behind children, educational performance"
wordcount         : "6,000"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

@@

classoption       : "man"
output            : papaja::apa6_pdf
appendix: 
  - "appendix.Rmd"

bibliography: references.bib
---

@@

## Introduction

@@

How are stay-behind children faring in the absence of parents? Research findings are mixed due to the differential effects of parental migration. Consider children's education. According to @venta2022 and @vanhook2020, labor migrants send remittances back, which can fund children's schooling as well as improve their nutritional and living conditions. To what degree can the inflow of remittances compensate for the loss of parental attention? Parental migration decreases the amount and quality of parental guidance, protection, and support. It is uncertain how these two opposing effects together shape children's educational performance. This uncertainty merits empirical investigation.

Quantitative social scientists have researched extensively about the education of left-behind children, but recent literature still has shortcomings. Some studies fail to capture the dynamic nature of migratory process as they draw upon cross-sectional data that focus on single time points [e.g. @ahmed2023; @vietnguyen2023; @gong2023]. Some other studies only probe correlation using regression analysis [e.g. @chen2023a; @liu2023; @huang2020]. These studies could have employed causal inference techniques.

@@

### China's Context

@@

The institutional discrimination against labor migrants finds its roots in China's "multilevel citizenship" system. As @vortherms2023 explained, the constituents of citizenship -- "membership rules" and "rights entitlements" -- operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through *hukou*, or the household registration system. The system differentiates people into local versus external and urban versus rural. In the rural-urban divide, urban hukou holders enjoy greater access to higher-quality welfare benefits and public services compared to rural hukou holders. In the local-external divide, local hukou holders are those registered in the same county where they live. In theory, Chinese citizens can migrate freely within the country. In practice, they cannot change hukou freely. When rural residents move to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the host city [@zhang2018; @chen2023c]. These institutions turn rural migrants into second-class citizens in effect. Migrant workers have organized petitions, strikes, and protests. These actions often fell on deaf ears and sometimes even ended with police crackdown [@chan2019; @göbel2021]. Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns. Only during the winter and summer breaks can these children reunite with parents.

How does China counter the issue of labor migration and left-behind children? According to @dong2020, the government has gradually liberalized the multilevel citizenship scheme. It has introduced residence permit for internal migrants as the basis for social entitlements. First-tier cities like Beijing and Shanghai have brought forth a point-based system to grant hukou to newcomers. The system's evaluation criteria include in-demand talents and skills, educational and professional qualifications, and consecutive years of employment and tax-payment. This parallels how European states differentially incorporate foreign immigrants into national welfare regimes [@pasquali2022]. Another policy response, as stated by @roberts2018, is the centralized boarding school. The Ministry of Education promotes boarding schools in the rural county-seats. These schools serve as a cost-effective way to enroll students in surrounding villages and to look after the left-behind children. In reality, such schools "are often ill-equipped to ensure the healthy upbringing and psychological wellbeing of children" (p. 6).

### Theory and Hypotheses

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources that can help children acquire human capital -- the stock of "knowledge, skills, and health that people accumulate over their lives" [@worldbank2020, p.1]. @yang2008 noted that in the Philippines, a sharp depreciation of peso increased the amount of remittances from migrants working abroad. This subsequently boosted children's school enrollment and educational spending. @vikram2021 performed propensity score matching and fixed-effect regression on a two-wave panel dataset featuring Indian children aged eight to eleven. She showed that fathers' migration was linked with increases in left-behind children's reading capability and educational expenditure (p. 991). In China, @gu2022a interviewed migrant workers and their family members. She found that they aspire to move up the social ladder; Such aspiration translates into "intensive and all-out investment in children's education" (p. 526). Oftentimes, labor migration presented the only way to sustain their children's education and livelihood. @peng2018 interviewed first-generation migrant mothers whose initial migration dated back to the 1990s. To them, "good mothering" primarily meant offering economic support for their children in education, which can help their children climb the ladder of success.

The other theoretical model is the family disruption model [@shen2021]. Children flourish under parental nurturing care. This term refers to the "stable environments that promote health and adequate nutrition, protect from threats ... opportunities for learning," and "responsive, emotionally supportive and developmentally enriching relationships" [@black2021, p. 1]. According to @chen2023, caregiver-child interactions are infrequent and unstable for left-behind children. The loss in such interactions "mediates almost all of the adverse effects of parental absence." This conclusion was based on structural equation modeling using a cross-sectional survey with a sample of Chinese children aged three to fifteen. @murphy2022 talked to children left behind in a Chinese village. Some of them "embraced the 'left-behind' label" because it captured their need for more parental care (p. 185), "especially companionship and affection in daily life" (p. 186). In Romania, @udrea2022 and @diaconu-gherasim2023 interviewed left-behind adolescents. Many of them felt disorientated about the lack of parents' support and oversight in schoolwork.

Given the two opposing effects, researchers have obtained mixed results on how parental migration impacts children's education. @botezat2020 adopted past immigration rate as an instrumental variable. They found a positive effect of parents' migration on academic grades of adolescents aged eleven to fifteen in Romania. @wang2019 collected data from junior high school students in central China. Their two-way fixed-effect model showed no significant effect of parental migration on left-behind children's achievement in mathematical test. @chae2019 analyzed the Migration and Remittances Household Survey of Kenya, Senegal, and Burkina Faso. They observed significant correlation between parental migration and left-behind children's school enrollment in Kenya, but not in the other two West African countries. @nguyen2016 conducted fixed-effect regression and demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight.

@@

Researchers have studied whether the impact differs between mother's migration and father's migration. @liu2023 analyzed how fifteen-year-old students' academic performance is associated with parental absence across 59 countries, using the 2012 Program of International Student Assessment data. Compared to students living with both parents, students perform worse in academic tests when only mother is absent or both parents are absent; When only father is absent, however, students are not significantly below par. @xu2019 adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had "little or no association with negative outcomes" on children's academic and cognitive performances. Only-mother absence, however, showed a "strong association with negative outcomes" (p. 1646). @chen2019 conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that maternal migration was "negatively associated with children's social competence and academic performance," while paternal migration had no direct effects on the two outcomes (pp. 860-861).

@murphy2020 and @fan2023 noted that in some Asian and Latin American societies, the social norms designate men as breadwinners and women as caregivers. Under such norms, fathers' migration is more likely to win children's acceptance than does mothers' migration. Left-behind children may even resent their migrant mothers for transgressing on the expected care-giving roles. I propose the second hypothesis:

@@

## Methods

### Sampling

@@

### Measurement

@@

The lists of covariates are in Table 1 -- Table 5 in the Results section. At the children's level, I controlled for age, sex, self-rated health, ethnicity, hukou type, number of siblings, whether the child had attended preschool, and whether the child attends boarding school. Age, sex, and health are associated with educational performance [@gu2021a]. Regarding ethnicity and hukou status, Han Chinese and urban hukou holders outperform ethnic minorities and rural hukou holders at school [@hasmath2023; @xu2018; @huang2020]. Attending preschool in early childhood and enrolling in boarding school in adolescence are predictors of educational performance [@liu2020a; @guo2021; @zheng2021]. I also included the number of siblings as a covariate, considering the quantity-quality trade-off in multi-child family [@doepke2015]. At the household level, I controlled for the levels of education of father and mother, whether the parent has a white collar job, whether the household has internet access, the perceived household financial conditions, and the number of extracurricular books at home. These variables measure family socioeconomic status, which is related to children's educational achievement [@liu2020; @young2018]. Finally, I controlled for the level of education that the child aspires to achieve and the parental expectation on the child's level of education. Children's academic achievement, their educational aspiration, and their parents' expectations are interrelated [@chen2023b; @xu2021]. All control variables were measured in the baseline survey.

### Analytical Strategy

@@

I adopted `R` (version 4.3.2) for all the analyses [@R-base]. This involved the following packages: `tidyverse` (version 2.0.0) and `haven` (version 2.5.3) for data cleaning [@R-tidyverse; @R-haven]; `fixest` (version 0.11.1) for econometric modeling [@R-fixest]; `MatchIt` (version 4.5.5) and `cobalt` (version 4.5.1) for matching [@ho2011; @R-MatchIt; @R-cobalt]; `modelsummary` (version 1.4.3), `vtable` (version 1.4.5), and `kableExtra` (version 1.3.4) for presenting the results [@R-modelsummary; @R-vtable; @R-kableExtra].

## Results

### Descriptive Statistics

Table 1 illustrates the independent variables, namely the migration status reported in the wave two survey. All `r printnum(nrow(d_cl))` children were living with parents in one household in baseline, but about sixteen percent saw at least one parent migrated between the two waves of survey. I call these students "newly left-behind children" and their households "new migration households." In these new migration households, only-mother migration and only-father migration each accounted for two out of five cases. The remaining one-fifth were both-parents migration.

\renewcommand{\arraystretch}{0.8}

```{r tab-idpvar}
tab_idpvar |>
  kbl(booktabs = T, caption = 'Independent variables by parental migration')
```

The dependent variables are summarized by the control group versus the treatment group in Table 2 and Figure 1 -- Figure 2. Compared to non-left-behind peers, newly left-behind children scored lower in wave two's cognitive and academic tests; Their cognitive test score increased less and their academic exam score dropped more. T-tests show that for the baseline outcomes, the differences of the mean values between-group are probably due to chance. For the wave two outcomes and the between-wave changes, the differences of the mean value are statistically significant.

```{r tab-dpvar}
tab_dpvar |> 
  kbl(booktabs = T, caption = 'Dependent variables by parental migration', row.names = F) |>
  add_indent(c(2,3,5,6)) |>
  add_header_above(c(" ", "No = 3529" = 2, "Yes = 675" = 2)) |>
  add_header_above(c(" ", "Parental Migration" = 4)) |>
  footnote(general = '* p < 0.05, ** p < 0.01, *** p < 0.001')
```

Table 3 summarizes the bivariate analysis on the binary control variables. New migration households had lower internet access rate. Among the newly left-behind children, male sex, past grade retention, and boarding at school were more common compared to the non-left-behind peers. Chi-square tests show that these differences in variance are unlikely due to chance.

```{r tab-ctvar-dit}
tab_ctvar_dit |>
  kbl(booktabs = T, caption = 'Dichotomous control variables by parental migration', row.names = F) |>
  add_header_above(c(" ", "No = 3529" = 2, "Yes = 675" = 2)) |>
  add_header_above(c(" ", "Parental Migration" = 4)) |>
  footnote(general = '0 = No, 1 = Yes; * p < 0.05, ** p < 0.01, *** p < 0.001')
```

Table 4 displays the bivariate analysis for categorical control variables, and Table 5 documents their measurement scales. While the majority of mothers and fathers had only finished high school or below, they held big expectations for children. Eight out of ten parents wished that their children complete university education. Scholars like @luo2013 and @ng2020 explained this pattern with the Chinese social norms and cultural traditions. In ancient China, the Confucian teaching emphasized the cultivation of personhood through learning, diligence, and persistence. In real life, knowledge could translate into power and privilege. Ordinary people could study for the highly competitive *keju* (civil service examination). As the world's "earliest meritocratic institution," this national exam system recruited top performers into the imperial government and the elite class [@chen2020, p. 2030]. Nowadays in China, the emperor and *keju* are long gone, but the influences are long-lasting. Chinese parents hope that their children receive the best education to secure a decent career [@luo2013; @ng2020].

```{r tab-ctvar-cnt}
tab_ctvar_cnt |>
  kbl(booktabs = T, caption = 'Categorical and continuous control variables by parental migration', row.names = F) |>
  add_header_above(c(" ", "No = 3529" = 2, "Yes = 675" = 2)) |>
  add_header_above(c(" ", "Parental Migration" = 4)) |>
  footnote(general = '* p < 0.05, ** p < 0.01, *** p < 0.001')
```

```{r tab-ctvar-msr}
tab_var |>
  kbl(booktabs = T, caption = 'Measurement and distribution of categorical variables')
```

### Regression Analysis

I first explain the results on children's cognitive test score. The unadjusted DID models show that newly left-behind adolescents are more likely to score lower in cognitive tests than do their non-left-behind counterparts (see Table 6). This lends support to Hypothesis 1a. Parental migration is linked to a `r coef(did_cog)[1] |> abs() |> round(2)` points decrease in cognitive test score. Hypothesis 2a is only partially supported. Compared with the non-left-behind peers, children with only father absent are associated with greater decrease (`r coef(did_cog_f_only)[1] |> abs() |> round(2)` points) in cognitive test score than those with only mother absent (`r coef(did_cog_m_only)[1] |> abs() |> round(2)` points). The adjusted DID models produce results similar to the unadjusted models, with one exception (see Table 7). That is, for the cognitive score, the both-parent-migrated coefficient's magnitude is lower than that of father-migrated.

@@

```{r tab-reg-cog-adj}
table_adj_cog
```

Next, I set forth the results on children's academic test grades. In line with Hypothesis 1b, the unadjusted DID models show that newly left-behind adolescents tend to perform worse in school exams than their non-left-behind peers (see Table 8 below). Parental migration is associated with a drop of `r coef(did_ttl)[1] |> abs() |> round(2)` points in academic exam score. Hypothesis 2b finds partial support. The exam grade of only-mother-migrated children sees a `r coef(did_ttl_m_only)[1] |> abs() |> round(2)` points decrease relative to non-left-behind counterparts. In the scenarios of only father migrated and both parents migrated, the results are not statistically significant. The adjusted DID models' results largely resemble those from the unadjusted models (See Table 9 in the appendix). Adding control variables measured in the baseline survey does not change the coefficients for parental migration.

@@

```{r tab-reg-ttl-adj}
table_adj_ttl
```

Some predictors are strong in both models of cognitive and academic abilities. These are children's age, preschool attendance, and educational aspiration as well as parents' expectation on children. For students, one year older in age is associated with decreases of `r coef(did_adj_cog)[3] |> abs() |> round(2)` points for cognitive test and of `r coef(did_adj_ttl)[3] |> abs() |> round(2)` points for academic exam. Preschool attendance is linked to increases of `r round(coef(did_adj_cog)[8],2)` points for cognitive test and of `r round(coef(did_adj_ttl)[8],2)` points for academic exam. Higher educational expectation from parents and greater educational aspiration from children are both linked to higher scores. The coefficients are `r round(coef(did_adj_cog)[14],2)` points and `r round(coef(did_adj_cog)[12],2)` points for cognitive test, respectively; and `r round(coef(did_adj_ttl)[14],2)` points and `r round(coef(did_adj_ttl)[12],2)` points for academic exams, respectively. Some predictors are strong in predicting one outcome but not the other. The covariate that only relates to academic grades is children's gender. Girls outperform boys by `r round(coef(did_adj_ttl)[4],2)` points in academic exams, but do not show advantage in cognitive tests. Another covariate, the number of siblings, i sonly relate to cognitive scores. Having an additional sibling is associated with lower cognitive score of `r coef(did_adj_cog)[7] |> abs() |> round(2)` points.

### Matching

Parental migration is not a randomly assigned treatment. Possibly, smart and ambitious students perform better at school, and high performance and ambition prompt parents to migrate to support children's schooling. Another scenario is that poverty may hinder children's educational achievement while driving parents to migrate. The endogeneity leads to selection bias. To tackle this bias, an applicable technique is matching. It aims to make two groups comparable by pairing a unit in the treatment group with one or several units in the control group based on observable covariates.

A popular choice for matching, as @pan2018 introduced, is propensity score matching (PSM). For each unit in the sample, PSM will first calculate a propensity score, namely the probability of receiving treatment based on the specified covariates. This score serves as a measure of the distance between two units. When handling this distance, PSM can match the units with two approaches. One is the nearest-neighbor matching. It examines all the treated units and selects the control unit closest in distance for pairing. This approach does not try to optimize any condition; each pair of units ignores how other pairing will occur or have occurred. The second approach is the optimal approach. It minimizes the total distance in propensity scores between the treatment and control groups. Optimal matching can assign a treated unit with more than one control unit, thereby retaining the full sample [@hansen2023].

@@

```{r tab-reg-m-adj-cog}
table_m_adj_cog
```

@@

```{r tab-reg-m-adj-ttl}
table_m_adj_ttl
```

## Discussion and Conclusion

@@

### Limitations

@@

### Recommendations

@@

# (APPENDIX) Appendix {.unnumbered}

@@
